### 2020-01-01 Chap-advlinmodels.Rmd
Setup for v 0.7.0

# Fortgeschrittene Themen der linearen Regression {#advlin}